---
title: "Who I Am and What I Want to Do: Linking Self-Construals and Career Interests"
---





Gore & Incera (2024)





#SETTING UP THE ANALYSES



Packages
```{r}
library(readxl)
library(performance)
library(tidyr)
library(lmboot)
```



Import data
```{r}
Dataset1 <- read_excel("Data1.xlsx")
Dataset1$Sex <- as.factor(Dataset1$Sex)
DatasetClean <- drop_na(Dataset1)
```






MODELS



CreativeArts



Model (Bimodal)
```{r}
m.CreativeArts = lm(CreativeArts ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.CreativeArts.Normality <- check_model(m.CreativeArts,
                                check = "normality")
m.CreativeArts.Normality
```

```{r}
m.CreativeArts.Boot <- wild.boot(CreativeArts ~ Sex + Independent + Interdependent + Relational + Physical, 
          B = 1000, data = DatasetClean, seed = NULL, bootDistn = "normal")

m.CreativeArts.Boot$origEstParam
```



PerformingArts



Model (Positively Skewed)
```{r}
m.PerformingArts = lm(PerformingArts ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.PerformingArts.Normality <- check_model(m.PerformingArts,
                                check = "normality")
m.PerformingArts.Normality
```

```{r}
m.PerformingArts.Boot <- wild.boot(CreativeArts ~ Sex + Independent + Interdependent + Relational + Physical, 
          B = 1000, data = DatasetClean, seed = NULL, bootDistn = "normal")

m.PerformingArts.Boot$origEstParam
```



Law



Model (Bimodal)
```{r}
m.Law = lm(Law ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Law.Normality <- check_model(m.Law,
                                check = "normality")
m.Law.Normality
```



Politics



Model  (Positively Skewed)
```{r}
m.Politics = lm(Politics ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Politics.Normality <- check_model(m.Politics,
                                check = "normality")
m.Politics.Normality
```



TechnicalWriting



Model (Positively Skewed)
```{r}
m.TechnicalWriting = lm(TechnicalWriting ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.TechnicalWriting.Normality <- check_model(m.TechnicalWriting,
                                check = "normality")
m.TechnicalWriting.Normality
```



InformationTech



Model  (Positively Skewed)
```{r}
m.InformationTech = lm(InformationTech ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.InformationTech.Normality <- check_model(m.InformationTech,
                                check = "normality")
m.InformationTech.Normality
```



Mathematics



Model  (Positively Skewed)
```{r}
m.Mathematics = lm(Mathematics ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Mathematics.Normality <- check_model(m.Mathematics,
                                check = "normality")
m.Mathematics.Normality
```



OfficeWork



Model (Bimodal)
```{r}
m.OfficeWork = lm(OfficeWork ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.OfficeWork.Normality <- check_model(m.OfficeWork,
                                check = "normality")
m.OfficeWork.Normality
```



Business



Model  (Positively Skewed)
```{r}
m.Business = lm(Business ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Business.Normality <- check_model(m.Business,
                                check = "normality")
m.Business.Normality
```



Management



Model  (Positively Skewed)
```{r}
m.Management = lm(Management ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Management.Normality <- check_model(m.Management,
                                check = "normality")
m.Management.Normality
```



Finance



Model  (Positively Skewed)
```{r}
m.Finance = lm(Finance ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Finance.Normality <- check_model(m.Finance,
                                check = "normality")
m.Finance.Normality
```



Sales



Model  (Positively Skewed)
```{r}
m.Sales = lm(Sales ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Sales.Normality <- check_model(m.Sales,
                                check = "normality")
m.Sales.Normality
```



FamilyActivity



Model  (Negatively Skewed)
```{r}
m.FamilyActivity = lm(FamilyActivity ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.FamilyActivity.Normality <- check_model(m.FamilyActivity,
                                check = "normality")
m.FamilyActivity.Normality
```



SocialServices



Model  (Negatively Skewed)
```{r}
m.SocialServices = lm(SocialServices ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.SocialServices.Normality <- check_model(m.SocialServices,
                                check = "normality")
m.SocialServices.Normality
```



OutdoorAgriculture



Model (Positively Skewed)
```{r}
m.OutdoorAgriculture = lm(OutdoorAgriculture ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.OutdoorAgriculture.Normality <- check_model(m.OutdoorAgriculture,
                                check = "normality")
m.OutdoorAgriculture.Normality
```



ManualLabor



Model  (Positively Skewed)
```{r}
m.ManualLabor = lm(ManualLabor ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.ManualLabor.Normality <- check_model(m.ManualLabor,
                                check = "normality")
m.ManualLabor.Normality
```



Engineering



Model (Positively Skewed)
```{r}
m.Engineering = lm(Engineering ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.Engineering.Normality <- check_model(m.Engineering,
                                check = "normality")
m.Engineering.Normality
```



ProtectiveServices



Model  (Positively Skewed)
```{r}
m.ProtectiveServices = lm(ProtectiveServices ~ Sex + Independent + Interdependent + Relational + Physical,
                data = Dataset1)

m.ProtectiveServices.Normality <- check_model(m.ProtectiveServices,
                                check = "normality")
m.ProtectiveServices.Normality
```










If you have questions, suggestions, and/or comments about the analysis, please contact Sara Incera at sara.incera@eku.edu





