---
title: "Wild Bootstrap Confidence Intervals"
format:
  lumo-html: 
    github-repo: "https://www.github.com/homerhanumat/wildboot"
    primary-color: "#a31f37"
    self-contained: true
    is-particlejs-enabled: true
author: Homer White
date: last-modified
---



## Questions

You had these three questions:

1. Do I need to change the `bootDIstn = "normal"` option to something else?
1. Can I run this bootstrap method if the distribution of the residuals is bimodal?
1. How can I get the confidence intervals for these five estimates? 

My answers:

1. No.  The options for `bootDIstn` implement various attempts to get slightly more powerful results at small sample sizes in specific situations. They do not correspond to the observed distribution of the residuals. 
1. Yes.  The bootstrap is designed to handle situations where one cannot make assumptions about the distribution of the residuals.
1. See the rest of this note.

But first let's attach some packages:

```{r setup}
#| warning: false
#| message: false
#| code-fold: false
library(tidyverse)
library(lmboot)
library(DT)
library(knitr)
library(readxl)
```

## A Small Example

From the `mtcars` data table in the **dataset** package:

```{r}
#| code-fold: false
mtcars2 <-
  mtcars %>% 
  mutate(
    vs = factor(vs), 
    am = factor(am)
  ) %>% 
  mutate(
    vs = fct_recode(vs, "V-shaped" = "0", "straight" = "1"),
    am = fct_recode(am, "automatic" = "0", "manual" = "1")
  ) %>% 
  select(mpg, disp, wt, vs, am)
```

Here's the example table:


```{r}
#| label: motor-trend
#| tbl-cap: "Motor Trend cars"
DT::datatable(
  mtcars2
)
```



